---
title: "Lab 10"
author: "Robert Saldivar"
date: "December 4, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Lab 10 Objectives:

- Correlation matrices
- Binary logistic regression
- Dates/times with lubridate
- Heatmap of solar irradiation in Santa Barbara, Hilo (HI), and Fairbanks (AK) in 2010
- Aquaculture with gganimate

library(gganimate) # For adding animation to ggplot graphics
library(ggimage) # For updating graph components with images

# PART 0. Load libraries
```{r packages}

library(tidyverse) # The tidyverse!
library(corrplot) # For correlation matrices
library(janitor) # For cleaning up column names
library(lubridate) # For dealing with dates & times



```

# PART 1. Correlation matrix (World Bank world environmental factors data)

Compiled World Bank data, accessed from: https://www.kaggle.com/zanderventer/environmental-variables-for-world-countries#World_countries_env_vars.csv

- Read in the "world_env_vars.csv file" as env_var
- Find correlations between all variables, columns 2:28
- Plot the correlation matrix (corrplot)

```{r env_corr}

env_var <- read_csv("world_env_vars.csv") %>%
  na.omit

cor_df <- cor(env_var[2:28])

corrplot(cor_df, 
         type = "upper",
         method = "ellipse",
         tl.col = "black",
         tl.cex = 0.5)

```

# PART 2. Binary Logistic Regression (Donner Party Data)

Use the 'glm' function for fitting *generalized linear models* (the *logit* - log odds of survival, in our case, will be linearly related to Sex and Age. So we expect the final model to look something like this: 

$$Log Odds (Survival) = \beta_0 + \beta_1(Age) + \beta_2(Sex)$$

We'll use 'family = binomial' to run binomial logistic regression...otherwise, this looks very similar to other types of regression we've already done. 

a. Read in the DonnerTable.csv file as DonnerTable

```{r donner_data}

DonnerTable <- read_csv("DonnerTable.csv")

```

b. Binomial logistic regression
1 = Survival
2 = Death

```{r blr}

donner_blr <- glm(Outcome ~ Sex + Age, family = "binomial", data = DonnerTable)
summary(donner_blr)

```

c. Use the model to answer: What are the log odds of survival for a 5 year-old female? The probability of survival?

```{r predict_f_5}

# 1. Create a data frame with variables Sex and Age, containing data "Female" and 5, respectively: 

f_5  <- data.frame(Sex = "Female", Age = 5)

# 2. Find the log odds of survival for the new data (5 year old female) using predict() function with type = "link":

f_5_logodds <- predict(donner_blr, newdata= f_5, type = "link")

# 3. Exponentiate the log odds to find ODDS of survival for a 5 year old female:

exp(f_5_logodds)

# Ask: Could we manually find the probability of survival for a 5-year old female? recall: p/(1-p) = ODDS

# 4. Actually, let's just use type = "response" in the predict function, which converts to a probability for us:

f_5_prob <- predict(donner_blr, newdata = f_5, type = "response")

```

d. What is the probability of survival for a 25 year-old male? 

```{r predict_m_25}

# Similarly:

m_25 <- data.frame(Sex = "Male", Age = 25) # Make a new data frame

m_25_prob <- predict(donner_blr, newdata = m_25, type = "response") # Find probability of survival
m_25_prob

```

e. Create new sequences of data so that we can graph probabilities for the entire spectrum of ages, designated by sex.

```{r new_data}

seq_age <- rep(seq(from = 0, to = 100), 2) # Create a sequence from 0 to 100, twice (one will be "Male" and one will be "Female")

f_101 <- rep("Female", 101) # Repeat 'Female' 101 times (to match years data)
m_101 <- rep("Male", 101) # Repeat 'Male' 101 times
mf_101 <- c(f_101, m_101) # Combine them into a single vector

# Combine the age and sex sequences into a single data frame - that will be the new data that we have our model make predictions for

donner_newdata <- data.frame(seq_age, mf_101) # MUST make column names match variables in the model!
colnames(donner_newdata) <- c("Age","Sex")

```

f. Now that we have new data to put into our model to have it make predictions, let's go ahead and actually find the predicted probabilities for each Age/Sex combination.

```{r donner_predict}

# Find probabilities using predict (with type = "response"). Include SE.

predicted_probs <- predict(donner_blr, newdata = donner_newdata, type = "response", se.fit = TRUE)

# Coerce outcome into data frame. 

graph_data <- data.frame(donner_newdata, predicted_probs$fit, predicted_probs$se.fit)

colnames(graph_data) <- c("Age", "Sex", "Probability", "SE")

```

g. Graph results.

```{r donner_graph}
ggplot(graph_data, aes(x = Age, y = Probability)) +
  geom_line(aes(color=Sex)) +
  geom_ribbon(aes(ymin=Probability - SE, ymax = Probability + SE, fill =Sex), alpha = 0.5)
  
```


# PART 3. Solar irradiation at the 3 locations in 2010 (pull in all together, do some lubridate stuff, etc.)

a. Read in multiple solar irradiation files (for SB (CA), Hilo (HI), and Fairbanks (AK)):
```{r solar_data}

si_full <- list.files(pattern = "solar_irradiation_*") %>%
  map_df(~read_csv(.)) %>%
  clean_names()


```

b. Wrangle the data
```{r wrangle_solar}

solar_tidy <- si_full %>%
  rename(
    sol_rad = etr_wh_m_2,
    date = yyyy_mm_dd,
    time = hh_mm_lst
  ) %>%
  filter(time !="NA") %>%
  mutate(site = fct_relevel(site , "Hawaii", "Santa Barbara", "Alaska"))

```

c. Use lubridate() functions to convert to times/dates
```{r date_time}

solar_tidy$date <- mdy(solar_tidy$date)
solar_tidy$time <- hms(solar_tidy$time)

```

d. Make an awesome figure of solar irradiation (heat/tile map)
```{r solar_graph}

solar_gg <- ggplot(solar_tidy, aes(x = date, y = time)) +
  geom_tile(aes(fill = sol_rad)) +
  scale_y_time() +
  facet_grid(site ~ .) +
  scale_fill_gradientn(colors=c("royalblue2", "mediumorchid1", "orange", "yellow")) 

solar_gg
  
```


#PART 4. gganimate example: total aquaculture production (metric tons) for United States, Brazil, Chile, and Ecuador

a. Get the data, and do some wrangling:
```{r aqua_data}
aq_df <- read_csv("aq_wb.csv")

aq_tidy <- aq_df %>%
  filter(country == "Brazil" | country == "Chile" | country == "Ecuador" |country == "United States") %>%
  gather(year, aq_prod, '1960':'2016') %>%
  filter(year >= 1990) %>%
  mutate(aq_mil = aq_prod/1000000) %>%
  select(country, year, aq_mil)


```


b. Read in the fish.png as 'fish'

```{r fish_image}

fish <- "fish.png"

```

c. Make a graph...with gganimate!
```{r gganimate_fish}

aq_plot <- ggplot(aq_tidy, aes(x = as.numeric(year), y = aq_mil, group = country)) +
  geom_line(aes(color=country)) +
  geom_point(aes(color = country)) +
  geom_text(aes(label = country, color = country), position = position_nudge(y = 0.04, x=1), size = 5)

aq_plot
```

##END LAB